Slitasje og kjørespor

### 2023-01-17 DRAFT_slitasjeOgKjørespor.Rmd
feat: Import HIA

#### Infrastructure index

The infrastructure index is explained
[here](https://brage.nina.no/nina-xmlui/handle/11250/2983607). I use it
as a kind of pressure or human influence index. It is likely to be a
good predictor for *slitasje*, and I think to use it to define local reference area,
or homogeneous (impact) areas.

```{r}
# The original data is here, on the R server, but it is not always working, so I also include a copy on the P server.

#path <- ifelse(dir == "C:", 
#        "R:/GeoSpatialData/Utility_governmentalServices/Norway_Infrastructure_Index/Original",
#        "/data/R/GeoSpatialData/Utility_governmentalServices/Norway_Infrastructure_Index/Original")
#

path <- ifelse(dir == "C:", 
        "P:/41201785_okologisk_tilstand_2022_2023/data/infrastrukturindeks",
        "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/infrastrukturindeks")

```

Import a stars proxy (no data imported yet)

```{r}
infra <- stars::read_stars(paste0(path, "/NY_INFRA_IND.tif"))
```

It's easier to see what's happening if we zoom in a bit. Lets get a
boundary box around Trondheim.

```{r}
myBB <- st_bbox(c(xmin=260520.12, xmax = 278587.56,
                ymin = 7032142.5, ymax = 7045245.27),
                crs = st_crs(infra))
```

Cropping the raster to speed up processing.

```{r}
infra_trd <- sf::st_crop(infra, myBB)
```

Transform to lat long due to osm requirements

```{r, warning=F}
infra_trd_ll <- sf::st_transform(infra_trd, 4326)
```

Get the boundary box of the cropped raster

```{r}
myBB_ll <- sf::st_bbox(infra_trd_ll)
```

Download highways using the above bbox.

```{r, warning=FALSE, message=F, eval=F}
hw <- 
  osmplotr::extract_osm_objects(
    bbox = myBB_ll,
    key = "highway",
    sf = T)
```

Transforming highway data back into utm, although not strictly
necessary.

```{r, eval=F}
hw_utm <- sf::st_transform(hw, sf::st_crs(infra_trd)) 
```

This object contains too many roads. I'll take out the unnamed roads.

```{r, eval=F}
hw_utm <- hw_utm[!is.na(hw_utm$name),]
```

Caching the sf object

```{r, eval=F}
saveRDS(hw_utm, "data/cache/highways_trondheim.rds")
```

Load object

```{r}
hw_utm <- readRDS("data/cache/highways_trondheim.rds")
```

```{r, fig.cap="Infrastrcture indeks over Trondheim. Major roads are in white."}
map_trd <- tm_shape(infra_trd)+
  tm_raster(title="Infrastructure Index",
            style="cont",
            palette = "viridis")+
  tm_layout(legend.outside = T)+
  tm_shape(hw_utm)+
  tm_lines(col="white")
map_trd
```

The infrastructure index is in EPSG32633. I will need to transform it.
It takes a couple of minutes, so I will store the result in P and load
it back in afterwards.

```{r, eval=F}
path <- ifelse(dir == "C:", 
        "P:/41201785_okologisk_tilstand_2022_2023/data/infrastrukturindeks",
        "/data/P-Prosjekter2/41201785_okologisk_tilstand_2022_2023/data/infrastrukturindeks")

infra <- sf::st_transform(infra, 25833)
stars::write_stars(infra, paste0(path, "/infrastrukturindeksEPSG_EPSG35833.tif"))
```

```{r}
infra2 <- stars::read_stars(paste0(path, "/infrastrukturindeksEPSG_EPSG35833.tif"))
```

Here I define a simplified categorical typology for the infrastructure
index using four classes.

```{r}
names(infra_trd) <- "infrastructureIndex"
infra_trd_reclassed <-  infra_trd %>%
  mutate(infrastructureIndex = case_when(
    infrastructureIndex < 1 ~ 0,
    infrastructureIndex < 6 ~ 1,
    infrastructureIndex < 9 ~ 2,
    infrastructureIndex >= 9 ~ 3
  ))
```

```{r}
map_trd_reclassed <- tm_shape(infra_trd_reclassed)+
  tm_raster(title="Infrastructure Index",
            #palette = "viridis",
            style="cat")+
  tm_layout(legend.outside = T)+
  tm_shape(hw_utm)+
  tm_lines(col="white")

```

```{r, fig.cap="Infrastrcture indeks over Trondheim, comparing the continous scale with the ordinal four-step scale. Major roads are in white."}
tmap_arrange(map_trd,
             map_trd_reclassed,
             ncol=1)
```

I tweaked the thresholds so that most of the forest next to Trondheim
was in the second lowest class. This looks pretty good to me.

@@

### Homogeneous impact areas

```{r, eval=F}
eaTools::ea_homogeneous_area(infra2)
```

@@

### Homogeneous areas
Export a map of the homogeneous areas that we aggregate the indicator
over:

```{r}
stars::write_stars(infra_trd_reclassed,
                   "data/homogeneous_areas/infrastructureIndex_fourClasses.tif")
```

>>>>

@@

### Homogeneous impact areas
I want to use the [homogeneous impact areas]() (HIA) to define smaller regions into which I can extrapolate the indicator values.
This data is generated by discretizing the Norwegian Infrastructure Index. I refer to the ordinal values of the
four impact area classes as their *Human Impact Factor*.

```{r}
HIA <- readRDS("P:/41201785_okologisk_tilstand_2022_2023/data/infrastrukturindeks/homogeneous_impact_areas.rds")
```

```{r, fig.cap="Counting the number of gridd cells for each Human Impact Factor."}
barplot(table(HIA$infrastructureIndex),
        ylab = "Number of 1x1 km grid cells",
        xlab = "Human impact factor")
```

I also want to split these four HIA classes based on their region.

<!-- Add function to ea_homogeneous_area() to find intersection with a decond delinimator -->

### Aggregate and spread (extrapolate)
I now want to find the mean indicator value for each HIA (i.e. to aggregate) and to spread these out
spatialy to the entire HIA within each region (i.e. to extrapolate).

```{r}
eaTools::ea_spread()
```

@@
